Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.

### 2024-02-28 docs/_tex/index.tex
part of the discuss

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.},
  pdfauthor={Simon Oiry; Bede Ffinian Rowe Davies; Pierre Gernez; Ana I. Sousa; Philippe Rosa; Maria Laura Zoffoli; Guillaume Brunier; Laurent Barillé},
  pdfkeywords={Drone, Remote Sensing},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

Remote sensing drone acquisitions are presented as a tool that can
potentially fill gaps left by satellite and in situ data. Drone can
cover large expanses while recording imagery at significantly higher
spatial resolutions than satellite (pixel size from cm to mm) and still
capturing data at multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping the spatial
distribution of species \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, many studies showcase their findings with
study case limited to a single flight, restricting the generalizability
of their application to other sites \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. This study aimed
to analyze the potential of a drone equipped with a multispectral sensor
for maping intertidal macrophytes, with a particular focus on
discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were
performed over soft-bottom intertidal areas along two European countries
(France and Portugal), covering a wide range of habitats, from
monospecific seagrass meadows to meadows mixed with green or red algae.
A deep learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without a
loss of accuracy in predictions.

@@

\subsection{Seagrass Discrimination}\label{seagrass-discrimination}

\subsection{Spatial and Spectral Resolution impact on the
prediction}\label{spatial-and-spectral-resolution-impact-on-the-prediction}

@@

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.},
  pdfauthor={Simon Oiry; Bede Ffinian Rowe Davies; Ana I. Sousa; Philippe Rosa; Maria Laura Zoffoli; Guillaume Brunier; Pierre Gernez; Laurent Barillé},
  pdfkeywords={Drone, Remote Sensing},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

Remote sensing drone acquisitions are presented as a tool that can
potentially fill gaps left by satellite and in situ data. Drone can
cover large expanses while recording imagery at significantly higher
spatial resolutions than satellite (pixel size from cm to mm) and still
capturing data at multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping the spatial
distribution of species \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, many studies showcase their findings with
study case limited to a single flight, restricting the generalizability
of their application to other sites \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. This study aimed
to analyze the potential of a drone equipped with a multispectral sensor
for mapping intertidal macrophytes, with a particular focus on
discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were
performed over soft-bottom intertidal areas along two European countries
(France and Portugal), covering a wide range of habitats, from
monospecific seagrass meadows to meadows mixed with green or red algae.
A deep learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without a
loss of accuracy in predictions.

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes on intertidal mudflats,
specifically focusing on distinguishing between Chlorophyceae (green
algae) and marine Magnoliopsida (seagrasses) using multispectral drone
data. The challenge arises from the shared pigment composition between
these two vegetation classes, complicating their differentiation using
remote sensing techniques. This challenge has been addressed through the
utilization of high-resolution drone flights allowing a precise
photointerpretation of pure pixels of the different vegetation classes
used as training data, providing a robust foundation for subsequent
classification analyses. In achieving 94.26 accuracy across all study
sites, this method demonstrates its efficacy in accurately classifying
macrophytes, thus offering a reliable approach for vegetation
classification in intertidal mudflats using remote sensing data.

\subsection{Green Macrophytes
Discrimination}\label{green-macrophytes-discrimination}

@@

The discrimination of seagrasses from green macroalgae presents
significant challenges, primarily due to the shared pigment composition
between these taxa, as depicted in Figure~\ref{fig-Pigm} but also by the
frequent spatial mixing of these green macrophytes. This similarity in
pigment composition, including photosynthetic and carotenoid pigments
like chlorophyll b and c, fucoxanthin, and neoxanthin, results in
spectral signatures that are often indistinguishable to non expererences
This spectral overlap is further complicated by the intermingling of
seagrasses and green macroalgae in the same spatial locations, posing a
significant challenge for remote sensing aimed at accurately mapping and
monitoring coastal ecosystems. Our study addresses these complexities by
utilizing high-resolution drone imagery with 10 spectral bands.
\citep{Davies2023} has shown that having at least 8 spectral bands
ranging between 500 nm to 850 nm with on band at 530 nm and another one
at 730 nm is crucial to accurately discriminate green macrophytes.

\phantomsection\label{cell-fig-ValidationGreen}
\begin{figure}[H]

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Validation/ConfusionMatrixGreen.png}

\caption{\label{fig-ValidationGreen}Sample of
Figure~\ref{fig-Validation} focusing on green macrophytes. The labels
inside the matrice indicate the number of pixels.}

The Micasense RedEdge-MX DUAL camera used in this study meets these two
criteria, enabling the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if the
pigment compostion of green macrophytes is similar, differences in the
spectral shape can still be observed (Figure~\ref{fig-vegetation}).
Several factors can explain these differences such has different pigment
concentration en proportion, different cellular organisation and 3D
disposition of the plant a whole on the intertidal mudflat
\citep[\citep{kirk1994light},
\citep{hedley2018influence}]{beach1997vivo}.

The VIP analysis of the Neural Network model (see Fig. VIP) shows that
the 531 nm band is the most important spectral band for accurately
identifying Chlorophyceae. In fact, at this wavelength, Chlorophyceae
exhibits the highest reflectance among all other classes, highlighting
the difference in carotenoid proportion between seagrasses and green
algae.

\subsection{Spectral and Temporal Resolution impact on the
prediction}\label{spectral-and-temporal-resolution-impact-on-the-prediction}

### 2024-02-28 docs/index.qmd
part of the discuss

@@

Remote sensing drone acquisitions are presented as a tool that can potentially fill gaps left by satellite and in situ data. Drone can cover large expanses while recording imagery at significantly higher spatial resolutions than satellite (pixel size from cm to mm) and still capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping the spatial distribution of species [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, many studies showcase their findings with study case limited to a single flight, restricting the generalizability of their application to other sites [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. This study aimed to analyze the potential of a drone equipped with a multispectral sensor for maping intertidal macrophytes, with a particular focus on discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green or red algae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of accuracy in predictions.

@@

@@

## Seagrass Discrimination

## Spatial and Spectral Resolution impact on the prediction

@@

>>>>

editor_options: 
  chunk_output_type: console
csl: RSE.csl

@@

Remote sensing drone acquisitions are presented as a tool that can potentially fill gaps left by satellite and in situ data. Drone can cover large expanses while recording imagery at significantly higher spatial resolutions than satellite (pixel size from cm to mm) and still capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping the spatial distribution of species [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, many studies showcase their findings with study case limited to a single flight, restricting the generalizability of their application to other sites [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. This study aimed to analyze the potential of a drone equipped with a multispectral sensor for mapping intertidal macrophytes, with a particular focus on discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green or red algae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of accuracy in predictions.

@@

######### Focus on green macrophytes
df_green<-df %>% 
  filter(Pred_Class %in% c("Magnoliopsida", "Chlorophyceae"),
         Truth %in% c("Magnoliopsida", "Chlorophyceae")) %>% 
  mutate(Pred_Class = as.factor(as.character(Pred_Class)),
         Truth = as.factor(as.character(Truth)))

validation_metrics_green<-confusionMatrix(df_green$Pred_Class,df_green$Truth)

conf_mat_green<-validation_metrics_green[[2]] %>% 
  as.numeric() %>%
  base::matrix(ncol = 2) %>%
  as.data.frame()

for(i in 1:(nrow(conf_mat_green))){
  if(i == 1 ){
    df_sum<-data.frame(row = NA,
                       col = NA)
  }
  df_sum[i,1] <- sum(conf_mat_green[i,],na.rm = T)
  df_sum[i,2] <- sum(conf_mat_green[,i],na.rm = T)
}
conf_mat_std_green<-conf_mat_green

for(row in 1:nrow(conf_mat_std_green)){
  for(col in 1:ncol(conf_mat_std_green)){
    conf_mat_std_green[row,col]<-conf_mat_std_green[row,col]/((df_sum$row[row] + df_sum$col[col])-conf_mat_std_green[row,col])
    }
}

confmat_rast_green<-conf_mat_std_green %>% 
  as.matrix() %>% 
  as.array() %>% 
  rast()

confmat_rast_green<-subst(confmat_rast_green, NA, -99)
names(confmat_rast_green)<-"layer"
ext(confmat_rast_green)<-c(0,2,0,2)
confmat_rast_green_sf<-as.polygons(confmat_rast_green,aggregate=F) %>% 
  sf::st_as_sf()
confmat_rast_green_centroid<-st_centroid(confmat_rast_green_sf)

values(confmat_rast_green)[values(confmat_rast_green) == 0] = NA
values(confmat_rast_green)[values(confmat_rast_green) == -99] = NA

labels_green<-validation_metrics_green[[2]] %>% 
  as.numeric() %>%
  base::matrix(ncol = 1) %>%
  as.data.frame() %>% 
  rename(label = "V1") %>% 
  mutate(x = c(0.5,0.5,1.5,1.5),
         y = c(1.5,0.5,1.5,0.5))

plot_green<-ggplot()+
geom_spatraster(data = confmat_rast_green,
                  na.rm = TRUE
                  )+
  geom_sf(data = confmat_rast_green_sf,
          fill = "transparent")+
  scale_fill_gradientn(
    # colours = c("black","firebrick4","goldenrod","yellow","chartreuse3"),
    # colours = c("khaki1","goldenrod","firebrick3","chartreuse4"),
    colours = c("khaki1","firebrick4","firebrick4","firebrick4","#3e0b0b"),
    # colours = c("#7d0000","#00b300","#00b300","#00b300","#007d00"),
    na.value = "white", 
    trans = "sqrt",
    limits = c(0,1),
    labels = c("0.05","0.25", "0.50", "1"),
    breaks = c(0.05,0.25,0.50,1))+
  # scale_fill_Bede("mixed", 
  #                 discrete = F,
  #                 reverse = TRUE,
  #                 na.value = "white",
  #                 trans = "sqrt",
  #                 labels = c("5","25", "50", "75", "100"),
  #                 breaks = c(0.05,0.25,0.50,0.75,0.9999))+
  labs(fill = "Proportion \n of pixels")+
  scale_x_continuous(breaks = seq(0.5,1.5, by = 1),
                     labels = c("Chlorophyceae","Magnoliopsida"),
                     expand = c(0, 0),
                     position = "top",
                     limits = c(0,2))+
  scale_y_continuous(breaks = seq(0.5,1.5, by = 1),
                     labels = c("Magnoliopsida","Chlorophyceae"),
                     expand = c(0, 0),
                     limits = c(0,2))+
  geom_label(data = labels_green,aes(x = x, y = y, label = label ))+
  xlab("References")+
  ylab("Prediction")+
  # ggplot2::annotation_custom(ggplotGrob(small_validation_matrices), xmin = 12, ymin = 0)+
  theme_Bede()+
  theme( 
        panel.border = element_blank(),
        axis.line=element_blank(),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20, angle =90, hjust = 0.5),

        # legend.position = "top",
        # legend.position = c(0.73, 0.9),
        # legend.direction="horizontal",
        # legend.key.width = unit(2,"cm"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20, vjust = 2),
        legend.title.align = 0.5,
        legend.justification = "center")

ggsave("./Figures/Validation/ConfusionMatrixGreen.png", plot_green, width = 671*4, height = 479*4, units = "px" )

@@

The primary objective of this study was to develop a method for the accurate classification of macrophytes on intertidal mudflats, specifically focusing on distinguishing between Chlorophyceae (green algae) and marine Magnoliopsida (seagrasses) using multispectral drone data. The challenge arises from the shared pigment composition between these two vegetation classes, complicating their differentiation using remote sensing techniques. This challenge has been addressed through the utilization of high-resolution drone flights allowing a precise photointerpretation of pure pixels of the different vegetation classes used as training data, providing a robust foundation for subsequent classification analyses. In achieving `r round(validation_metrics$overall[[1]]*100,2)` accuracy across all study sites, this method demonstrates its efficacy in accurately classifying macrophytes, thus offering a reliable approach for vegetation classification in intertidal mudflats using remote sensing data.

## Green Macrophytes Discrimination

@@

The discrimination of seagrasses from green macroalgae presents significant challenges, primarily due to the shared pigment composition between these taxa, as depicted in @fig-Pigm but also by the frequent spatial mixing of these green macrophytes. This similarity in pigment composition, including photosynthetic and carotenoid pigments like chlorophyll b and c, fucoxanthin, and neoxanthin, results in spectral signatures that are often indistinguishable to non expererences This spectral overlap is further complicated by the intermingling of seagrasses and green macroalgae in the same spatial locations, posing a significant challenge for remote sensing aimed at accurately mapping and monitoring coastal ecosystems. Our study addresses these complexities by utilizing high-resolution drone imagery with 10 spectral bands. @Davies2023 has shown that having at least 8 spectral bands ranging between 500 nm to 850 nm with on band at 530 nm and another one at 730 nm is crucial to accurately discriminate green macrophytes.

::::: columns
::: column
```{r Global Accuracy assesment Plot}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: Sample of @fig-Validation focusing on green macrophytes. The labels inside the matrice indicate the number of pixels.
#| label: fig-ValidationGreen
#| fig-width: 10
#| fig-height: 10
#| out-width: "100%"

knitr::include_graphics("./Figures/Validation/ConfusionMatrixGreen.png")

```
:::

::: column
The Micasense RedEdge-MX DUAL camera used in this study meets these two criteria, enabling the classifier to achieve `r round(validation_metrics_green$overall[[1]],2)*100`% of accuracy between these two classes (@fig-ValidationGreen). Even if the pigment compostion of green macrophytes is similar, differences in the spectral shape can still be observed (@fig-vegetation). Several factors can explain these differences such has different pigment concentration en proportion, different cellular organisation and 3D disposition of the plant a whole on the intertidal mudflat [@beach1997vivo, @kirk1994light, @hedley2018influence]. 
:::
:::::

The VIP analysis of the Neural Network model (see Fig. VIP) shows that the 531 nm band is the most important spectral band for accurately identifying Chlorophyceae. In fact, at this wavelength, Chlorophyceae exhibits the highest reflectance among all other classes, highlighting the difference in carotenoid proportion between seagrasses and green algae.

## Spectral and Temporal Resolution impact on the prediction
